---
title: "Predicting Flood Severity of Northwest Florida Counties using Random Forest Regression Modeling - Spring 2026 "
author: "Meagan Russell (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---
<!--

This is for me later, once the slides are ready to go.

Slides: [slides.html](slides.html){target="_blank"} ( Go to `slides.qmd`
to edit)
--!>

## Week 1 - Individual Research:

Paper One: Random Forest Algorithm Overview

This paper provides an overview of the Random Forest (RF) algorithm. Random Forest is a type of machine learning algorithm that uses a target variable known as a dependent variable. The goal of Random Forest is to predict the dependent variable based on independent variables. The way RF works is by creating decision trees. Decision trees are a series of yes/no questions, for example: 


Did the flooding cause any deaths? 

Did the flooding result in more than $1,000,000 in property damage?

Did the flood occur countywide?

Did the flood cause agricultural damage?


The authors go on to explain how Random Forest is a smart model that uses simple models, decision trees, which each can make small mistakes, but in the end, once all of the simple models are combined, they get the answer right. In addition, each decision tree's selection of data is random, which prevents the model from memorizing the data. There are limitations to Random Forest, such as increased computational cost and reduced transparency; however, overall Random Forest is reliable, widely applicable, and suitable for real-world problems.

Cite new paper [@Salman2024RandomForest]

```{r, warning=FALSE, echo=T, message=FALSE}
# loading packages 
library(tidyverse)
library(knitr)
library(ggthemes)
library(ggrepel)
library(dslabs)
```

```{python}
import pandas as pd
```

```{r, warning=FALSE, echo=TRUE}
# Load Data
kable(head(murders))

ggplot1 = murders %>% ggplot(mapping = aes(x=population/10^6, y=total)) 

  ggplot1 + geom_point(aes(col=region), size = 4) +
  geom_text_repel(aes(label=abb)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(formula = "y~x", method=lm,se = F)+
  xlab("Populations in millions (log10 scale)") + 
  ylab("Total number of murders (log10 scale)") +
  ggtitle("US Gun Murders in 2010") +
  scale_color_discrete(name = "Region")+
      theme_bw()
  

```

### Modeling and Results

-   Explain your data preprocessing and cleaning steps.

-   Present your key findings in a clear and concise manner.

-   Use visuals to support your claims.

-   **Tell a story about what the data reveals.**

```{r}

```

### Conclusion

-   Summarize your key findings.

-   Discuss the implications of your results.

## References
